---
title: "BiomassScript"
output:
  html_document: default
  html_notebook: default
---

This script is written to provide a quick check of biomass harvest data from Baxters block.  
There are 2 sections:  
1. General correlations/checks  
2. Treatment means  

***

```{r Reading data, include=FALSE}
# load readxl library
library(readxl)

# read column headers
header <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", skip = 12, col_names = FALSE, n_max = 1)

# read data so we get correct col types (without interference from units)
harvestData <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", skip = 14, col_names = FALSE)

# combine headers and data
names(harvestData) <- header

# print data and summarise
print(harvestData)
str(harvestData)

# change data types ## add to this as I go...not sure if character or factor are better in this case
harvestData$Plot <- as.character(harvestData$Plot)
harvestData$Rep <- as.factor(harvestData$Rep)
harvestData$Irr_Num <- as.character(harvestData$Irr_Num)
harvestData$Zone_ID <- as.factor(harvestData$Zone_ID)

summary(harvestData)
names(harvestData)

## remove redundant columns???

# store column headers with units in tibble for later reference
headerUnits <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", n_max = 1, 
                            skip = 12, col_names = TRUE)
```

```{r, include=FALSE}
# select harvest no. and display date 
library(dplyr)
harvestData <- harvestData  %>%
  filter(Harvest_Num == "H3") ### CHANGE THIS FILTER FOR EACH HARVEST 
```

Data displayed is from `r harvestData$Harvest_Num[1]`, harvested on `r harvestData$Harvest_Date[1]`.  

***  
<br>
<br>
  
## Correlations and checks  

```{r libraries and levels for correlations, include=FALSE}
library(ggplot2)
library(car)
library(gridExtra)

# set order of Zone_ID
levels(harvestData$Zone_ID)
harvestData$Zone_ID <- factor(harvestData$Zone_ID, levels = c("Low", "Mid", "High"))
levels(harvestData$Zone_ID)
```

<br>
  
### Number of samples  

```{r, echo=FALSE}
# checking number of samples
ggplot(data = harvestData, aes(Rep)) +
  geom_bar(aes(colour=Irr_Num)) +
  facet_grid(~Zone_ID) +
  scale_y_continuous(breaks = seq(0, 10, 1)) + # making sure dimensions are available in case data wrong 
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```


<br>
  
### Calculation checks  

```{r, echo=FALSE}
# calculations - expecting 1:1 relationships
plot1 <- ggplot(data = harvestData, aes(x = Biomass_Yield_Harvest_Moist, y = Whole_Sample_FW)) +
  geom_point() +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
harvestData$Sum_Component_Yields <- (harvestData$Green_Leaf_Yield + harvestData$Stem_Yield + harvestData$Dead_Leaf_Yield + harvestData$Ear_Yield)
plot2 <- ggplot(data = harvestData, aes(x = Total_Biomass_Dry, y = Sum_Component_Yields)) +
  geom_point() +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
grid.arrange(plot1, plot2, ncol = 2)
```


<br>
  
### Yield component correlations  

```{r, echo=FALSE}
# yield components
scatterplotMatrix(harvestData[c("Total_Biomass_Dry", "Green_Leaf_Yield", "Stem_Yield", 
                                "Dead_Leaf_Yield", "Ear_Yield")],
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright")
```

<br>
  
### Leaf area correlations  

```{r, echo=FALSE}
# leaf area
scatterplotMatrix(harvestData[c("Whole_Sample_FW", "Green_leaf_area", "LAI", 
                                "Green_Leaf_Yield")],
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright")

```

<br>
  
### Green leaf correlations  

```{r, echo=FALSE}
# green leaf
scatterplotMatrix(harvestData[c("Whole_Sample_FW",  "Leaf_FWT", "Green_leaf_DW",
                                 "Green_Leaf_Yield", "Total_Biomass_Dry")], 
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright")
```

<br>
  
### Dead leaf correlations  

```{r, echo=FALSE}
# dead leaf
scatterplotMatrix(harvestData[c("Whole_Sample_FW",  "Dead_leaf_FW", "Dead_leaf_DW",               
                                "Dead_Leaf_Yield", "Total_Biomass_Dry")], 
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright")
```

<br>
  
### Stem correlations  

```{r, echo=FALSE}
# stem
scatterplotMatrix(harvestData[c("Whole_Sample_FW",  "Stem_FW", "Stem_DW", "Stem_Yield",
                                "Total_Biomass_Dry")], 
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright")

```

<br>
  
### Ear correlations  

```{r, echo=FALSE}
# ear
scatterplotMatrix(harvestData[c("Whole_Sample_FW",  "Ear_FW", "Ear_DW", "Ear_Yield",
                                "Total_Biomass_Dry")], 
                  diagonal = "none",  smoother = FALSE, reg.line = FALSE,
                  groups = harvestData$Zone_ID, by.groups = TRUE,
                  legend.pos = "bottomright") 
```


<br>
<br>
  
## Treatment means  

```{r libraries and levels for treatment means, include=FALSE}
library(ggplot2)
library(dplyr)

# set order of Zone_ID
levels(harvestData$Zone_ID)
harvestData$Zone_ID <- factor(harvestData$Zone_ID, levels = c("Low", "Mid", "High"))
levels(harvestData$Zone_ID)
```
<br>
  
### FW of samples (g) across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = Whole_Sample_FW)) + 
  geom_boxplot() + 
  geom_point(aes(colour = Irr_Num)) +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```
<br>
<br>
  
### Total DM (kg/ha) across zones and irrig 
  
#### Grouped by zone  
```{r, echo=FALSE}
harvestData %>% 
  group_by(Zone_ID) %>% 
  summarise(N=length(Total_Biomass_Dry),
      mean=mean(Total_Biomass_Dry),
      min=min(Total_Biomass_Dry),
      max=max(Total_Biomass_Dry),
      med=median(Total_Biomass_Dry),
      sd=sd(Total_Biomass_Dry),
      se=sd/sqrt(N))
ggplot(data = harvestData, aes(x = Zone_ID, y = Total_Biomass_Dry)) + 
  geom_boxplot() + 
  geom_point(aes(colour = Irr_Num)) +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

#### Grouped by irrig  
```{r, echo=FALSE}
harvestData %>% 
  group_by(Irr_Num) %>% 
  summarise(N=length(Total_Biomass_Dry),
      mean=mean(Total_Biomass_Dry),
      min=min(Total_Biomass_Dry),
      max=max(Total_Biomass_Dry),
      med=median(Total_Biomass_Dry),
      sd=sd(Total_Biomass_Dry),
      se=sd/sqrt(N))
ggplot(data = harvestData, aes(x = Irr_Num, y = Total_Biomass_Dry)) + 
  geom_boxplot() + 
  geom_point(aes(colour = Zone_ID)) +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

#### Grouped by zone and irrig  
```{r, echo=FALSE}
TotalBiomasssummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(Total_Biomass_Dry),
      mean=mean(Total_Biomass_Dry),
      min=min(Total_Biomass_Dry),
      max=max(Total_Biomass_Dry),
      med=median(Total_Biomass_Dry),
      sd=sd(Total_Biomass_Dry),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(TotalBiomasssummary)
pd <- position_dodge(0.1)
ggplot(data = TotalBiomasssummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean total biomass") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```


<br>
<br>
  
### LAI across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = LAI)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num) +
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())

LAIsummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(LAI),
      mean=mean(LAI),
      min=min(LAI),
      max=max(LAI),
      med=median(LAI),
      sd=sd(LAI),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(LAIsummary)
pd <- position_dodge(0.1)
ggplot(data = LAIsummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean LAI") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

<br>
<br>
  
### Green leaf yield across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = Green_Leaf_Yield)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num) +
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())

GreenLeafsummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(Green_Leaf_Yield),
      mean=mean(Green_Leaf_Yield),
      min=min(Green_Leaf_Yield),
      max=max(Green_Leaf_Yield),
      med=median(Green_Leaf_Yield),
      sd=sd(Green_Leaf_Yield),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(GreenLeafsummary)
pd <- position_dodge(0.1)
ggplot(data = GreenLeafsummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean green leaf yield") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

<br>
<br>
  
### Dead leaf yield across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = Dead_Leaf_Yield)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num) +
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())

DeadLeafsummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(Dead_Leaf_Yield),
      mean=mean(Dead_Leaf_Yield),
      min=min(Dead_Leaf_Yield),
      max=max(Dead_Leaf_Yield),
      med=median(Dead_Leaf_Yield),
      sd=sd(Dead_Leaf_Yield),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(DeadLeafsummary)
pd <- position_dodge(0.1)
ggplot(data = DeadLeafsummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean dead leaf yield") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

<br>
<br>
  
### Stem yield across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = Stem_Yield)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num) +
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())

Stemsummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(Stem_Yield),
      mean=mean(Stem_Yield),
      min=min(Stem_Yield),
      max=max(Stem_Yield),
      med=median(Stem_Yield),
      sd=sd(Stem_Yield),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(Stemsummary)
pd <- position_dodge(0.1)
ggplot(data = Stemsummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean stem yield") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```

<br>
<br>
  
### Ear yield across zones and irrig  
```{r, echo=FALSE}
ggplot(data = harvestData, aes(x = Zone_ID, y = Ear_Yield)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num) +
  theme_bw() +
  theme(axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())

Earsummary <- harvestData %>% 
  group_by(Zone_ID, Irr_Num) %>% 
  summarise(N=length(Ear_Yield),
      mean=mean(Ear_Yield),
      min=min(Ear_Yield),
      max=max(Ear_Yield),
      med=median(Ear_Yield),
      sd=sd(Ear_Yield),
      se=sd/sqrt(N),
      ymax=mean+se,
      ymin=mean-se)
print(Earsummary)
pd <- position_dodge(0.1)
ggplot(data = Earsummary, aes(x = Zone_ID, y = mean)) + 
  geom_point(aes(colour = Irr_Num), position = pd) +
  geom_errorbar(aes(ymax = ymax, ymin = ymin, colour = Irr_Num), width = 0.1, position = pd) +
  labs(y = "mean ear yield") +
  theme_bw() +
  theme(panel.border=element_blank(),
        axis.line=element_line(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank())
```






```{r Extras for treatment means, eval=FALSE, include=FALSE}

# visualise no. of plots
ggplot(data = harvestData, aes(Rep)) +
  geom_bar(aes(colour=Irr_Num)) +
  facet_grid(~Zone_ID)


# total biomass boxplot with n=4
ggplot(data = harvestData, aes(x = Zone_ID, y = Total_Biomass_Dry)) + 
  geom_boxplot(aes(colour = Irr_Num))

# total biomass scatter plot version which may be better
ggplot(data = harvestData, aes(x = Zone_ID, y = Total_Biomass_Dry)) + 
  geom_point(aes(colour = Irr_Num)) +
  facet_grid(~ Irr_Num)


# LAI boxplot with n=4
ggplot(data = harvestData, aes(x = Zone_ID, y = LAI)) + 
  geom_boxplot(aes(colour = Irr_Num))

ggplot(data = harvestData, aes(x = Irr_Num, y = LAI)) + 
  geom_boxplot(aes(colour = Zone_ID))

# LAI scatter to add back in once there's expected STD vs VRI differences
ggplot(data = harvestData, aes(x = Irr_Num, y = LAI)) + 
  geom_point(aes(colour = Zone_ID)) +
  facet_grid(~ Zone_ID)

# LAI boxplot version that may be better than scatters?
ggplot(data = harvestData, aes(x = Zone_ID, y = LAI)) + 
  geom_boxplot() + 
  geom_point(aes(colour = Irr_Num))


```





```{r Scraps, eval=FALSE, include=FALSE}
str(harvestData)



library(readxl)
harvest <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", skip = 12, col_names = TRUE,
                      col_types = c("date", "text", "numeric", rep("text", 3), rep("numeric", 39)))

head(harvest)

harvest <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", skip = 12, col_names = TRUE)
str(harvest)


trial <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", col_names = FALSE)
head(trial, 20)


#read header and row of units

header <- sapply(read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", nrows = 2, 
                            skip = 12),paste,collapse="_")
result <- setNames(read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", skip = 14, header = FALSE), nm)

test <- read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx",skip=14)

head(result,2)

setNames( 1:3, c("foo", "bar", "baz") )

tmp <- 1:3
names(tmp) <-  c("foo", "bar", "baz")
tmp

ma <- matrix(1:12, 3, 4)
nrow(ma)   # 3
ncol(ma)   # 4

path <- readxl_example("geometry.xls")
read_excel(path)
read_excel(path, range = "B3:D6")
read_excel(path, range = "C3:D5")
read_excel(path, range = "A3:D5")
read_excel(path, range = "A4:E5")
read_excel(path, range = "C5:E7")
read_excel(path, range = anchored("C4", dim = c(3, 2)), col_names = FALSE)


header <- sapply(read_excel("K:/CPDiary/Data/BaxtersMVI/Biomass.xlsx", n_max = 2, 
                            skip = 12, col_names = FALSE),paste,collapse="_")

names(data) <- headers



# library(GGally)
# ggpairs(harvestData[c("Whole_Sample_FW", "Biomass_Yield_Harvest_Moist",
#                                 "Total_Biomass_Dry","Green_leaf_area", "LAI", 
#                                 "Green_Leaf_Yield", "Stem_Yield", "Dead_Leaf_Yield", "Ear_Yield")])

# ggpairs(harvestData[c("Whole_Sample_FW",  "Ear_FW", "Ear_DW", "Ear_Yield",
#                                 "Total_Biomass_Dry")], 
#        # mapping = aes(colour = Zone_ID)) # this part not working

# legend(10,20, pch=c(3,2), col=c("red", "blue","green"), c("Low", "Mid","High"), bty="o", cex=.8, box.col="darkgreen")


ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = Green_leaf_area)) +
  geom_point()
ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = LAI)) +
  geom_point()

ggplot(data = harvestData, aes(x = Green_leaf_area, y = LAI)) +
  geom_point()
ggplot(data = harvestData, aes(x = Partitioning_Total_FW, y = LAI)) +
  geom_point()
ggplot(data = harvestData, aes(x = Whole_Sample_FW, y = LAI)) +
  geom_point()
ggplot(data = harvestData, aes(x = Green_leaf_area, y = Leaf_FWT)) +
  geom_point()

ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = Total_Biomass_Dry)) +
  geom_point()
ggplot(data = harvestData, aes(x = LAI, y = Total_Biomass_Dry)) +
  geom_point()

coef(lm(Green_Leaf_Yield ~ Leaf_FWT, data = harvestData))
ggplot(data = harvestData, aes(x = Leaf_FWT, y = Green_Leaf_Yield)) +
  geom_point() +
  geom_abline(intercept = 1685.31494, slope = -37.63116)
ggplot(data = harvestData, aes(x = Green_leaf_DW, y = Green_Leaf_Yield)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
ggplot(data = harvestData, aes(x = Stem_FW, y = Stem_Yield)) +
  geom_point() 
ggplot(data = harvestData, aes(x = Stem_DW, y = Stem_Yield)) +
  geom_point()
ggplot(data = harvestData, aes(x = Dead_leaf_FW, y = Dead_Leaf_Yield)) +
  geom_point() 
ggplot(data = harvestData, aes(x = Dead_leaf_DW, y = Dead_Leaf_Yield)) +
  geom_point()
ggplot(data = harvestData, aes(x = Ear_FW, y = Ear_Yield)) +
  geom_point() 
ggplot(data = harvestData, aes(x = Ear_DW, y = Ear_Yield)) +
  geom_point() 

ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = Dead_Leaf_Yield)) +
  geom_point()
ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = Stem_Yield)) +
  geom_point()
ggplot(data = harvestData, aes(x = Green_Leaf_Yield, y = Ear_Yield)) +
  geom_point()
ggplot(data = harvestData, aes(x = Stem_Yield, y = Ear_Yield)) +
  geom_point()



ggplot(data = harvestData, aes(x = Biomass_Yield_Harvest_Moist, y = Total_Biomass_Dry)) +
  geom_point()

#checking calculations, expect exact 1:1 relationship
ggplot(data = harvestData, aes(x = Biomass_Yield_Harvest_Moist, y = Whole_Sample_FW)) +
  geom_point()

#additional column for summing component kgDM/ha to determine equivalence to Total_Biomass_Yield
harvestData$Sum_Component_Yields <- (harvestData$Green_Leaf_Yield + harvestData$Stem_Yield + harvestData$Dead_Leaf_Yield + harvestData$Ear_Yield)
ggplot(data = harvestData, aes(x = Total_Biomass_Dry, y = Sum_Component_Yields)) +
  geom_point() +
  geom_line()


```


